Exploring the evolution of stellar rotation using Galactic kinematics

### 2020-04-24 paper/ms63.tex
v. small changes

The velocity dispersions of stars in the \mct\ sample, shown in figure
\ref{fig:vplot}, provide the following picture of rotational evolution.
At young ages \citep[younger than around 1 Gyr but still old enough to be on
the main sequence and have transitioned from the `C' sequence to the `I'
sequence ][]{barnes2003}, stellar rotation period {\it decreases} with {\it
increasing} mass.
This is likely because lower-mass stars with deeper convection zones have
stronger magnetic fields, larger Alfv\'en radii and therefore experience
greater angular momentum loss rate \citep[\eg][]{schatzman1962, kraft1967,
parker1970, kawaler1988, charbonneau2010, matt2012, matt2015}.
% This aligns with the current assumptions about stellar spin-down, dynamo
% theory, and the gyrochronology paradigm that has been in place for decades.
% \citep[\eg][]{skumanich1972, noyes1984, kawaler1988, barnes2003, barnes2010}.
According to the \citet{spada2019} model, the radiative cores and convective
envelopes of stars are decoupled at these young ages, \ie\ transportation of
angular momentum from the surface to the core of the star is reduced, so the
surface slows down due to wind-braking but the core keeps spinning rapidly.
According to the data presented in figure \ref{fig:vplot}, at intermediate
ages, the rotation periods of K dwarfs appear {\it constant} with mass, and at
late ages rotation period {\it increases} with {\it increasing} mass.
The interpretation of this, according to the \citet{spada2019} model, is that
lower-mass stars are still braking more efficiently at these intermediate and
old ages \racomment{
    compared to higher-mass stars,
}
but their cores are more
tightly coupled to their envelopes, allowing angular momentum transport
between the two layers.
Angular momentum surfaces \racomment{
    from the core
}
and prevents the stellar
envelopes from spinning-down rapidly, and this effect is strongest for late
K-dwarfs with effective temperatures of $\sim$4000-4500K and masses
$\sim$0.5-0.7 M$_\odot$.

>>>>

The velocity dispersions of stars in the \mct\ sample, shown in figure
\ref{fig:vplot}, provide the following picture of rotational evolution.
\racomment{
    At young ages \citep[younger than around 1 Gyr but still old enough to be on
the main sequence and have transitioned from the `C' sequence to the `I'
sequence ][]{barnes2003}, stellar rotation period {\it decreases} with {\it
increasing} mass.
}
This is likely because lower-mass stars with deeper convection zones have
stronger magnetic fields, larger Alfv\'en radii and therefore experience
greater angular momentum loss rate \citep[\eg][]{schatzman1962, kraft1967,
parker1970, kawaler1988, charbonneau2010, matt2012, matt2015}.
% This aligns with the current assumptions about stellar spin-down, dynamo
% theory, and the gyrochronology paradigm that has been in place for decades.
% \citep[\eg][]{skumanich1972, noyes1984, kawaler1988, barnes2003, barnes2010}.
According to the \citet{spada2019} model, the radiative cores and convective
envelopes of stars are decoupled at these young ages, \ie\ transportation of
angular momentum from the surface to the core of the star is reduced, so the
surface slows down due to wind-braking but the core keeps spinning rapidly.
According to the data presented in figure \ref{fig:vplot}, at intermediate
ages, the rotation periods of K dwarfs appear {\it constant} with mass, and at
late ages rotation period {\it increases} with {\it increasing} mass.
The interpretation of this, according to the \citet{spada2019} model, is that
lower-mass stars are still braking more efficiently at these intermediate and
old ages \racomment{
    compared to higher-mass stars,
}
but their cores are more
tightly coupled to their envelopes, allowing angular momentum transport
between the two layers.
Angular momentum surfaces \racomment{
    from the core
}
and prevents the stellar
envelopes from spinning-down rapidly, and this effect is strongest for late
K-dwarfs with effective temperatures of $\sim$4000-4500K and masses
$\sim$0.5-0.7 M$_\odot$.
